Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-01-21 2_SSL_EES/article/article_supplementary.tex
updated supplementary information

\section{Manufacturing Cost Model}

@@

As with the flux metrics previously introduced, efficacy in lighting is likewise dependent on the luminosity function of figure \ref{fig:tristimulus}. A light source emitting very \textit{efficiently} in the infrared yet emitting no visible light has a very low \textit{efficacy}. The luminous efficacy of radiation $K$ is mathematically defined as the normalized, integrated product of the spectral radiant flux of a light source with the wavelength dependent human sensitivity to light \cite{cie-term-effrad}
%
\begin{equation}
\label{eqn:ler}
    K [\text{lm/W}_{opt}]= \frac{\int_0^\infty K( \lambda ) \phi \text{d} \lambda}{\int_0^\infty \phi \text{d} \lambda}
\end{equation}
%
where
%
\begin{align*}
    K &\dots \text{spectral luminous efficacy} \\
    \phi &\dots \text{spectral radiant flux} \\
    \lambda &\dots \text{wavelength}
\end{align*}
%
This metric can be computed from spectral data alone and does not require additional spectral normalization. It enables straightforward comparison between the performance of different downconversion phosphors, as shown in figure \ref{fig:se_efficacy} in chapter \ref{ch:results}. Light sources emitting in the far red or blue part of the spectrum have lower efficacy of radiation, as the product with the luminosity function of figure \ref{fig:tristimulus} shows. Care must be taken not to confuse this efficacy metric with efficacy of source, which depends not only on the spectral power distribution of a light source.

This metric describes the match of a light-emitting diode package spectrum to the human visual system. It thus describes only a part of the total device performance. It is used in the computation of other metric introduced in section \ref{subsec:subeff} below.

    The luminous efficacy of a light source $\eta$ is defined as the ratio between the emitted luminous flux and the consumed electrical power \cite{cie-term-effsrc}
    %
    \begin{equation}
        \eta [\text{lm/W}_{el}]= \frac{\phi}{P_{el}}
    \end{equation}
    %
    Unlike the luminous efficacy of radiation, this metric cannot be extracted from the spectrum of a light source. It is often cited in device datasheets, scientific literature and textbooks when describing the performance of light-emitting diodes. Care must be taken not to confuse this efficacy metric with efficacy of radiation, which depends only on the spectral power distribution of a light source.

    This metric fulfills the requirements laid out in section \ref{sec:metrics_requirements}. However, as it captures overall device efficacy, it depends on a large number of device properties and parameters. This makes attribution to single changes in device design or manufacturing difficult.

\subsection{Spectral Data of Identifies LED Phosphor Innovations}

\begin{figure}[h!]
	\centering
    \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/phosphor_spectrum-comparison.pdf}
	\caption{Two key metrics associated with phosphor innovations in light-emitting diodes and their associated spectral data. Top panel: Luminous efficacy of radiation (LER) and colour rendering index (CRI) of white LED devices represented in Main Article Figure 7 (Historical Improvements in Consumer Experience Metrics of Phosphor-Converted White LED). Years indicate the earliest identified representative white LED products with published spectral data that used phosphor innovations listed in Main Article Table 2. The desirable direction of improvements towards higher luminous efficacy at higher CRI is indicated by a red arrow. Metrics were calculated from spectral data shown in the two bottom rows using the \texttt{colour-science} package for Python \cite{colour-science_software}. Bottom two rows of panels: Corresponding spectral data. The luminosity function, describing the wavelength-dependent sensitivity of the human eye, is shown for reference in red in each panel. Note that peaks or large tails of the device emission spectrum at the far ends of the luminosity function are not desirable. This is because photons of the corresponding energy are lost to the human observer and count towards the spectral loss channel. Abbreviations: Arb. Units - Arbitrary Units (compare section \ref{}). Plot legends indicate the years of publication of the spectral data and phosphor mixtures used, with the following designations: $^*$ \ch{CaSrS:Eu^{2+}} $^\star\beta$-SiAlON:Eu$^{2+}$, $^\dagger$\ch{Lu_3Al_5O_{12}:Ce^{3+}}, $^\ddagger$\ch{(Ba,Sr)_2Si_5N_8:Eu^{2+}}. Full chemical formulas for displayed phosphors: \ch{Sr[Li2Al2O2N2]:Eu^{2+}} (SALON), \ch{Sr[LiAl_3N_4]:Eu^{2+}} (SLA), \ch{K2SiF6:Mn^{4+}} (PFS), \ch{Sr2Si5N8:Eu^{2+}} (258), Quantum Dots (QD). . Source (top panel): own elaboration based on spectral data. Source (spectral data): adapted from published spectral data for LEDs with the following phosphors: YAG \cite{bando1998development}, YGAG \cite{Mueller2002}, 258 \cite{MuellerMach2005}, SLA \cite{Pust2014}, PFS \cite{trigain_spectrum}, QD \cite{lumileds2016qd}\cite{osram2019qd}, SALON \cite{Hoerder2019}.}
\label{fig:phosphor_spectrum}
\end{figure}

@@

Prior to Nakamura’s invention of highly efficient blue LEDs in the 1990s, Japan’s Nichia Corporation had not sold commercially successful semiconductor products, instead specializing in phosphors for cathode ray tubes (CRT) and fluorescent lamps \cite{nakamura2013blue}  . Nevertheless, extensive firm expertise in this area helped Nichia’s Yoshinori Shimizu formulate the principles of using CRT phosphors to convert blue light from Nakamura’s LEDs into white light in 1994 \cite{shimizu1994sheet},\cite{cho2017white}. By 1996, Shimizu and his colleagues developed \cite{bando1996}, \cite{shimizu1999light} the first practical LED application of a well-known Yttrium Aluminium Garnet (YAG) CRT phosphor activated with cerium \cite{blasse1967new}, enabling the first commercial white LED products manufactured and sold by Nichia since late 1996 \cite{bando1998development}, \cite{cho2017white}.

Importantly, the YAG phosphor does not exhibit the spectral properties desirable in general illumination applications (see Figure 8 in the main text). An early solution to this problem, which was first discovered in the late 1967 \cite{holloway1969optical} and suggested for LEDs by the same team at Nichia in 1996 \cite{bando1998development},\cite{shimizu1999light}, was to use the gadolinium-doped red-shifted YAG phosphor (YGAG). Used in combination with red-emitting sulfide phosphors, by 2002 it helped bring to the market the first generation of warm white light LED products, e.g., those produced by Lumileds \cite{Mueller2002}. However, sulfide phosphors led to accelerated deterioration of sensitive parts of LED devices and became less efficient as operating temperatures increased. New chemically stable and non-toxic red phosphors were needed.

@@

The most recent red narrow-band phosphor breakthrough included in Table 3 and Figure 8 in the main text, indicated as SALON, has been under development during the late 2010s by a group of Austrian and German researchers that included Huppertz, the discoverer of the “258” material, working in collaboration with Osram, another major LED manufacturer \cite{seibald2019phosphor}, \cite{Hoerder2019} , \cite{Hoerder2020}. The first patent application for this phosphor was filed in 2016. The SALON phosphor is a derivate of the SLA phosphor. Therefore, it is the only innovation related to consumer experience metrics identified in our study that seemingly not involved technology spillovers.

@@

@@

@@

\begin{table}[!ht]
    \centering
    \begin{tabular}{|l|l|l|l|l|}
    \hline
        \# & sector & role & country & expertise \\ \hline
        1 & academia & sr. researcher & UK & epitaxy \\ \hline
        2 & industry & consultant, form. sr. researcher & USA & device architecture \\ \hline
        3 & industry & consultant, form. head of R\&D & Germany & epitaxy \\ \hline
        4 & academia & professor & Austria & phosphors \\ \hline
        5 & industry & consultant, form. head of R\&D & USA & device architecture \\ \hline
        6 & consulting & consultant, form. sr. technical advisor & USA & device architecture \\ \hline
        7 & academia & professor & Germany & phosphors \\ \hline
        8 & government & R\&D manager & USA & device architecture \\ \hline
        9 & consulting & consultant & USA & device applications \\ \hline
        10 & academia & professor & France & device physics \\ \hline
        11 & industry & sr scientist, form. head of R\&D & USA & device architecture \\ \hline
        12 & industry & principal scientist & Germany & phosphors \\ \hline
        13 & industry & form. head of R\&D & USA & phosphors \\ \hline
    \end{tabular}
    \label{tab:interviews}
    \caption{List of domain experts with whom semi-structured interviews were conducted. \\ Abbreviations: form. - former, sr. - senior}
\end{table}

>>>>

\section{Lighting Metrics}

Efficacy in lighting is dependent on the luminosity function, which describes the wavelength-dependent sensitivity of the human eye. A light source emitting very \textit{efficiently} in the infrared yet emitting no visible light has a very low \textit{efficacy}. The luminous efficacy of radiation $K$ is mathematically defined as the normalized, integrated product of the spectral radiant flux of a light source with the wavelength dependent human sensitivity to light \cite{cie-term-effrad}

\begin{equation}
\label{eqn:ler}
    K [\text{lm/W}_{opt}]= \frac{\int_0^\infty K( \lambda ) \phi \text{d} \lambda}{\int_0^\infty \phi \text{d} \lambda}
\end{equation}

\begin{align*}
    K &\dots \text{spectral luminous efficacy} \\
    \phi &\dots \text{spectral radiant flux} \\
    \lambda &\dots \text{wavelength}
\end{align*}

This metric can be computed from spectral data alone and does not require additional spectral normalization. It enables straightforward comparison between the performance of different downconversion phosphors, as shown in figure \ref{fig:se_efficacy} in chapter \ref{ch:results}. Light sources emitting in the far red or blue part of the spectrum have lower efficacy of radiation, as the product with the luminosity function of figure \ref{fig:tristimulus} shows. Care must be taken not to confuse this efficacy metric with efficacy of source, which depends not only on the spectral power distribution of a light source.

This metric describes the match of a light-emitting diode package spectrum to the human visual system. It thus describes only a part of the total device performance.

In contrast to the efficacy \textit{of radiation} $K$, the luminous efficacy \textit{of a light source} $\eta$ is defined as the ratio between the emitted luminous flux and the consumed electrical power \cite{cie-term-effsrc}

\begin{equation}
    \eta [\text{lm/W}_{el}]= \frac{\phi}{P_{el}}
\end{equation}

Unlike the luminous efficacy of radiation, this metric cannot be extracted from the spectrum of a light source. It is often cited in device datasheets, scientific literature and textbooks when describing the performance of light-emitting diodes. Care must be taken not to confuse this efficacy metric with efficacy of radiation, which depends only on the spectral power distribution of a light source. As it captures overall device efficacy, it depends on a large number of device properties and parameters. This makes attribution to single changes in device design or manufacturing difficult.

\subsection{Arbitrary Units}
\label{subsec:arbunit}

In spectroscopy, an arbitrary unit (abbreviated "Arb. Unit") is a relative unit of measurement to show the ratio of the intensity to a predetermined reference. In the case of Figure \ref{fig:phosphor_spectrum}, the reference is the highest point in the spectrum.

\newpage
\section{Manufacturing Cost Model}

@@

@@

Prior to Nakamura’s invention of highly efficient blue LEDs in the 1990s, Japan’s Nichia Corporation had not sold commercially successful semiconductor products, instead specializing in phosphors for cathode ray tubes (CRT) and fluorescent lamps \cite{nakamura2013blue}  . Nevertheless, extensive firm expertise in this area helped Nichia’s Yoshinori Shimizu formulate the principles of using CRT phosphors to convert blue light from Nakamura’s LEDs into white light in 1994 \cite{shimizu1994sheet}\cite{cho2017white}. By 1996, Shimizu and his colleagues developed \cite{bando1996}\cite{shimizu1999light} the first practical LED application of a well-known Yttrium Aluminium Garnet (YAG) CRT phosphor activated with cerium \cite{blasse1967new}, enabling the first commercial white LED products manufactured and sold by Nichia since late 1996 \cite{bando1998development}\cite{cho2017white}.

Importantly, the YAG phosphor does not exhibit the spectral properties desirable in general illumination applications (see Figure 8 in the main text). An early solution to this problem, which was first discovered in the late 1967 \cite{holloway1969optical} and suggested for LEDs by the same team at Nichia in 1996 \cite{bando1998development}\cite{shimizu1999light}, was to use the gadolinium-doped red-shifted YAG phosphor (YGAG). Used in combination with red-emitting sulfide phosphors, by 2002 it helped bring to the market the first generation of warm white light LED products, e.g., those produced by Lumileds \cite{Mueller2002}. However, sulfide phosphors led to accelerated deterioration of sensitive parts of LED devices and became less efficient as operating temperatures increased. New chemically stable and non-toxic red phosphors were needed.

@@

The most recent red narrow-band phosphor breakthrough included in Table 3 and Figure 8 in the main text, indicated as SALON, has been under development during the late 2010s by a group of Austrian and German researchers that included Huppertz, the discoverer of the “258” material, working in collaboration with Osram, another major LED manufacturer \cite{seibald2019phosphor}\cite{Hoerder2019}\cite{Hoerder2020}. The first patent application for this phosphor was filed in 2016. The SALON phosphor is a derivate of the SLA phosphor. Therefore, it is the only innovation related to consumer experience metrics identified in our study that seemingly not involved technology spillovers.

@@

Down conversion with ultra-narrow-band phosphor can achieve the highest spectral efficiency. However, few such phosphors have been identified, with even less exhibiting desirable material properties such as thermal stability \cite{Phillips2007}. The first commercially successful ultra-narrow-band red phosphor was developed by General Electric (GE). It is based on a potassium fluorosilicate (PFS) compound activated with manganese ions. Its luminescence was first recorded by Paulusz at GE in 1972 \cite{paulusz1973efficient}. In the early 2000s, while searching for potential new LED phosphor materials for GE’s lighting business at GE Lumination, Emil Radkov rediscovered Paulusz’s findings in the literature. Following extensive research on PFS chemical synthesis and material properties conducted in collaboration with the University of Sofia in Bulgaria, Radkov’s Alma Mater, the PFS phosphor had been under development at GE since 2005 \cite{radkov2006red}\cite{radkov2009red}. This work, supported by public funding from the U.S. Department of Energy (DOE) Solid-State Lighting program , resulted in a series of critical improvements in the PFS phosphor properties \cite{Setlur2010}\cite{lyons2012color}, eventually enabling its commercialization under the “TriGain” brand in 2015 \cite{trigain_spectrum}\cite{setlur2015trigain}\cite{Murphy2015}.

@@

Quantum dots (QD) are semiconductor nanocrystals whose quantum size effects make QDs behave as “artificial atoms”. Semiconductor quantum dots were first synthesized in the Soviet Union in 1981 \cite{ekimov1981quantum} and at Bell Labs in the U.S. in 1983 \cite{Rossetti1983}. Luminescent properties of quantum dots were first empirically observed in 1984 \cite{fojtik1984photo} and extensively studied in the early 1990s. The key feature of QD luminescence discovered in those studies is that its colour is determined by the QD particle size, making it possible to create pure monochromatic blue, green and red light sources just by tuning the QD size. The first application of QDs in LEDs was reported in 1994 in an electroluminescent hybrid QD-polymer LED. However, this LED type could not be used in general illumination due to its very low luminous efficacy. An alternative application of QDs as a kind of a “phosphor” for light down conversion from an LED light source was proposed in the early 2000s as part of the U.S. Department of Energy (DOE)-funded “A Revolution in Lighting“ project at Sandia National Laboratory \cite{simmonsfinal}. This concept was successfully demonstrated by Sandia researchers on a commercial LED in 2003 \cite{shea_rohwer_development_2004}\cite{noauthor_sandia_nodate} and was swiftly taken up and advanced further by a group in Taiwan \cite{chen_white-light_2005}\cite{hsueh-shih_chen_ingan-cdse-znse_2006} The first commercial application of QDs in an LED lamp was brought about by a collaboration between an MIT-born startup QD Vision and the U.S.-based luminaire manufacturer Nexxus Lighting in 2009 \cite{ledprof_nexxusqd}, \cite{bourzac2013quantum}. However, rapid advances in the spectral and conversion performance of down-conversion phosphors and high manufacturing cost of quantum dots resulted in the discontinuation of this product. After finding market success in display backlighting first demonstrated by Samsung in 2010 \cite{Jang2010} and commercialized by QD Vision in Sony television sets in 2013 \cite{bourzac2013quantum}, QDs returned to the general lighting market in products offered by Lumileds \cite{noauthor_global_2017}\cite{noauthor_quantum_2020} around 2017 and Osram in 2019 \cite{osramqdots} in the form of mid-power LED packages that combined QDs with traditional phosphors for light down conversion.

\subsection{Spectral Data of Identified LED Phosphor Innovations}

\begin{figure}[H]
	\centering
    \includegraphics[width=0.95\textwidth]{2_SSL_EES/article/figures/phosphor_spectrum-comparison.pdf}
	\caption{Two key metrics associated with red phosphor innovations in light-emitting diodes and their associated spectral data. Top panel: Luminous efficacy of radiation (LER) and colour rendering index (CRI) of white LED devices represented in Main Article Figure 7 (Historical Improvements in Consumer Experience Metrics of Phosphor-Converted White LED). Years indicate the earliest identified representative white LED products with published spectral data that used phosphor innovations listed in Main Article Table 2. The desirable direction of improvements towards higher luminous efficacy at higher CRI is indicated by a red arrow. Metrics were calculated from spectral data shown in the two bottom rows using the \texttt{colour-science} package for Python \cite{colour-science_software}. Bottom two rows of panels: Corresponding spectral data. The luminosity function, describing the wavelength-dependent sensitivity of the human eye, is shown for reference in red in each panel. Note that peaks or large tails of the device emission spectrum at the far ends of the luminosity function are not desirable. This is because photons of the corresponding energy are lost to the human observer and count towards the spectral loss channel. Abbreviations: Arb. Units - Arbitrary Units (compare section \ref{subsec:arbunit}). Plot legends indicate the years of publication of the spectral data and phosphor mixtures used with the following designations: ? -  other parts of phosphor mixture not disclosed, $^*$ - \ch{CaSrS:Eu^{2+}}, $^\star$ - $\beta-$SiAlON:Eu$^{2+}$, $^\dagger$ - \ch{Lu_3Al_5O_{12}:Ce^{3+}}, $^\ddagger$ - \ch{(Ba,Sr)_2Si_5N_8:Eu^{2+}}. Full chemical formulas for displayed phosphors: \ch{Sr[Li2Al2O2N2]:Eu^{2+}}(SALON), \ch{Sr[LiAl_3N_4]:Eu^{2+}} (SLA), \ch{K2SiF6:Mn^{4+}} (PFS), \ch{Sr2Si5N8:Eu^{2+}} (258), Quantum Dots (QD). Source (top panel): own elaboration based on spectral data. Source (spectral data): adapted from published spectral data for LEDs with the following phosphors: YAG \cite{bando1998development}, YGAG \cite{Mueller2002}, 258 \cite{MuellerMach2005}, SLA \cite{Pust2014}, PFS \cite{trigain_spectrum}, QD \cite{lumileds2016qd}\cite{osram2019qd}, SALON \cite{Hoerder2019}.}
\label{fig:phosphor_spectrum}
\end{figure}

@@

\begin{table}[h!]
\small
    \centering
    \begin{tabular}{|l|l|l|l|l|}
    \hline
        \textbf{\#} & \textbf{sector} & \textbf{role} & \textbf{country} & \textbf{expertise} \\ \hline
        1 & academia & sr. researcher & UK & epitaxy \\ \hline
        2 & industry & consultant, form. sr. researcher & USA & device architecture \\ \hline
        3 & industry & consultant, form. head of R\&D & Germany & epitaxy \\ \hline
        4 & academia & professor & Austria & phosphors \\ \hline
        5 & industry & consultant, form. head of R\&D & USA & device architecture \\ \hline
        6 & consulting & consultant, form. sr. technical advisor & USA & device architecture \\ \hline
        7 & academia & professor & Germany & phosphors \\ \hline
        8 & government & R\&D manager & USA & device architecture \\ \hline
        9 & consulting & consultant & USA & device applications \\ \hline
        10 & academia & professor & France & device physics \\ \hline
        11 & industry & sr scientist, form. head of R\&D & USA & device architecture \\ \hline
        12 & industry & principal scientist & Germany & phosphors \\ \hline
        13 & industry & form. head of R\&D & USA & phosphors \\ \hline
    \end{tabular}
    \label{tab:interviews}
    \caption{List of domain experts with whom semi-structured interviews were conducted. \\ Abbreviations: form. - former, sr. - senior}
\end{table}
